Evaluating infectious disease forecasts with allocation scoring rules

### 2023-12-01 alloscore_manuscript/alloscore-application.Rnw
updates to application reorg

<<score-computation>>=
tar_load("Kat15k_alloscores")

@@

tar_load("score_data")

@@

<<metrics-rankings, results = "asis">>=
## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
all_scores |>
  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
  select(model, alloscore, allo_rank, mwis, mwis_rank) |>
  arrange(allo_rank) |>
  xtable::xtable(label = "tab:one-week-results",
                 caption = "caption TBD") |>
  print(include.rownames = FALSE)
@

Forecasts and allocations for the nine states with the highest numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
\ngr{Aaron - can you add this file to the Sweave doc? Maybe just make a target for it so the figure is generated outside the latex?}
As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
A direct comparison between the forecasts from the JHUAPL-Bucky and USC-SI\_kJalpha models show that while the JHUAPL-Bucky forecast distributions were closer to the truth across all states, the allocations suggested by those forecasts often were more inefficient than those from the USC-SI\_kJalpha model (Figure \ref{fig:thermometer-plot}).

@@

\subsubsection{Forecast scores showed differences over time}
<<grob-hosps-over-time>>=
tar_load("truth_data")
dat_to_plot <- truth_data |>
  filter(location == "US",
         target_end_date %in% score_data_summaries$target_end_date)
p1 <- ggplot(dat_to_plot,
             aes(x=target_end_date, y=value)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=15000, linetype=2) +
  ylab("hospital admissions") +
  xlab(NULL) +
  ylim(c(0, NA))
@

>>>>

<<score-computation>>=
tar_load("Kat15k_alloscores")
tar_load("score_data")
tar_load("truth_data")
tar_load("forecast_data")

@@

@@

<<metrics-rankings, results = "asis">>=
## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
all_scores |>
  filter(target_end_date == "2022-01-10", !is.na(mwis)) |>
  select(model, alloscore, allo_rank, mwis, mwis_rank) |>
  arrange(allo_rank) |>
  xtable::xtable(label = "tab:one-week-results",
                 caption = "caption TBD") |>
  print(include.rownames = FALSE)
@

Forecasts and allocations for the nine states with the highest numbers of hospitalizations on January 3, 2022 reveal mechanics about how allocations are made (Figure \ref{fig:thermometer-plot}).
As described in the methods above, allocations for a given location from a specific model are generated by finding the single quantile of the probabilistic forecast across all locations that reaches the resource allocation limit (in this case, assumed to be 15,000).
A direct comparison between the forecasts from the JHUAPL-Bucky and USC-SI\_kJalpha models show that while the JHUAPL-Bucky forecast distributions were closer to the truth across all states, the allocations suggested by those forecasts often were more inefficient than those from the USC-SI\_kJalpha model (Figure \ref{fig:thermometer-plot}).
\ngr{add in some specific interpretations from the figure.}

<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location. The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
source("R/plot_functions.R")
tar_load(alloscore_2021.12.20)
plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
          start_date = "2021-12-01",
          stop_date = "2022-01-25",
          f_date = "2021-12-20",
          te_date = "2022-01-03",
          ## highest locations in terms of hospitalizations
          locations = c("FL", "NY", "TX", "CA"),
          f_width1 = 4, f_alpha = .4,
          allocations = TRUE,
          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
          alloscores = alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date")),
          truth = truth_data,
          f_colors = c("#4A708B", "#CD2626"),
          one_K = 15000) + labs(x = NULL)
@

@@

\subsubsection{Forecast scores showed differences over time}
<<grob-hosps-over-time>>=
dat_to_plot <- truth_data |>
  filter(location == "US",
         target_end_date %in% score_data_summaries$target_end_date)
p1 <- ggplot(dat_to_plot,
             aes(x=target_end_date, y=value)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=15000, linetype=2) +
  ylab("hospital admissions") +
  xlab(NULL) +
  ylim(c(0, NA))
@
